---
title: "Installation instructions"
author: "Lisa Rottjers"
date: "27 januari 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To install _manta_ and _anuran_, it is best to use [venv](https://docs.python.org/3/library/venv.html) to make sure you do not get conflicting dependencies. 
The installation instructions are slightly different for Windows and Unix systems. 

[Unix instructions](https://ramellose.github.io/networktutorials/installation.html#unix)

[Windows instructions](https://ramellose.github.io/networktutorials/installation.html#windows)

## Unix instructions {#unix}

To make sure you are creating a Python 3 virtual environment, run venv as follows. Replace the path with a path on your Unix system. The demo folder will contain the virtual environment. 

```{r unix1, eval=FALSE}
python3 -m venv /home/user/Documents/demo
```

It is possible that you cannot yet use venv. In that case, follow the instructions in the error message to download venv for your Unix system. 

Activate or deactivate the virtual environment:

```{r unix2, eval=FALSE}
source /home/user/Documents/demo/bin/activate
deactivate
```

If you activate the virtual environment, your terminal should say (demo) before your username. 
With the activated environment, download and install _manta_ and _anuran_. 

```{r unix3, eval=FALSE}
python3 -m pip install git+https://github.com/ramellose/manta.git
python3 -m pip install git+https://github.com/ramellose/anuran.git
```

Test if you can run the software by calling the help command. 

```{r unix4, eval=FALSE}
manta -h
anuran -h
```

## Windows instructions {#windows}

To make sure you are creating a Python 3 virtual environment, run venv as follows. Replace the path with a path on your Windows system. The demo folder will contain the virtual environment. 

```{r windows1, eval=FALSE}
python3 -m venv C:/user/Documents/demo
```

Activate or deactivate the virtual environment:

```{r windows2, eval=FALSE}
C:/user/Documents/demo/Scripts/activate.bat
deactivate
```

If you activate the virtual environment, your terminal should say (demo) before your username. 
With the activated environment, download and install _manta_ and _anuran_. 

```{r windows3, eval=FALSE}
python3 -m pip install git+https://github.com/ramellose/manta.git
python3 -m pip install git+https://github.com/ramellose/anuran.git
```

Test if you can run the software by calling the help command. 

```{r windows4, eval=FALSE}
manta -h
anuran -h
```
