---
title: "Microbiome Data Analyses Workshop"
author: "Sam Rottjers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This hands-on workshop is part of the Microbiome Data Analyses Workshop (Hasselt, 2021). 
In this workshop, we will introduce FlashWeave, a new method for network inference and we will walk you through the steps necessary to start analyzing your network with other software, including manta and anuran. Please find included links to installation instructions and all tutorials. 

We recommend first going through the FlashWeave tutorial, so you have a network to work with. 
However, after that, feel free to choose an analysis method of your choice to start with; these can all be followed independently. 

## Installation instructions
* [Install FlashWeave (and Julia)](https://ramellose.github.io/networktutorials/installation_flashweave.html)
* [Install Python 3](https://www.python.org/downloads/)
* [Install PyCharm or another Python editor of choice](https://www.jetbrains.com/pycharm/download/#section=windows)
* [Install manta](https://github.com/ramellose/manta)
* [Install anuran](https://github.com/ramellose/anuran)

### Installing Python packages
You can choose to install Python packages from PyCharm, but you can also do so from the command line. 
Use the command below to download networkx, pandas and biom-format. When you install manta and anuran, all other dependencies should be installed automatically. 
```{r packages, message=FALSE, eval=FALSE}
python3 -m pip install networkx
```

## Tutorials
* [Running FlashWeave on sponge networks](https://ramellose.github.io/networktutorials/running_flashweave.html) 
* [Working with igraph in R](https://ramellose.github.io/networktutorials/igraph_networks.html)  
* [Working with NetworkX in Python](https://ramellose.github.io/networktutorials/python_networks.html)  
* [Running manta on FlashWeave networks](https://ramellose.github.io/networktutorials/workshop_manta.html)  
* [Running anuran on FlashWeave networks](https://ramellose.github.io/networktutorials/workshop_anuran.html) 

## Glossary

Some of these terms may be used interchangably throughout the workshop, but they are not exactly the same. Please find included the definitions as I use them below. 

* Network: a collection of nodes and edges.  
* Graph: Synonymous to network. 
* Node: A piece of information in a network which usually represents a microbial taxon for microbial networks. 
* Edge: A link between two nodes. 
* Weight: A property of an edge that describes the strength of the relationship. Weights can often be negative or positive. 
* Association: A type of edge which does not represent an observed microbial interaction. 
* Interaction: An ecological behaviour that involves two or more organisms, where they affect each other in some way. 
* Directed: For a directed network, edges describe the direction of an effect. I.e. A affects B. 
* Undirected: For an undirected network, edges do not describe the direction of an effect. We can only conclude that A is associated to B. 
* Set: A collection of objects. Many sets are derived from specific collections, e.g. intersections or unions. 
* Intersection: An intersection is a set that contains overlap between one or more groups. 
* Union: A union is a complete collection, containing all objects belonging to all groups. 
* Difference: A difference is a set that contains objects only found in one group. 
* Partial intersection: A partial intersection is an intersection that contains overlap in a minimum number of groups. 

[Overview of all tutorials](https://ramellose.github.io/networktutorials/index.html)


