---
title: "Running FlashWeave on sponge networks"
author: "Lisa Rottjers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this tutorial, we will write a small script to run network inference with FlashWeave on a group of BIOM files. These files are derived from samples collected for the Sponge Microbiome Project (Moitinho-Silva et al., 2017). Each file contains a collection of samples collected from a single sponge order. 

You can download the processed files from the Github repository:
[link to zip BIOM files](https://github.com/ramellose/networktutorials/blob/master/Workshop%202021/sponges.zip).   Unzip the files in a location of your choice. 

In the script below, change the data_path `C:/Users/Username/Documents/Workshop/sponges` to the location where your files are stored. The script will read each of the files in the folder, so make sure only the BIOM files are in the folder. 

```{r flashweave, message=FALSE, eval=FALSE}
using FlashWeave

data_path = string("C:/Users/Username/Documents/Workshop/sponges")

files = readdir(data_path)
for file in files:
  file_path = string(data_path, '/', file)
 	network = learn_network(file_path, sensitive=true, heterogeneous=false, n_obs_min=10)
 	saveloc = string(data_path, '/', file[begin:end-5], ".gml")
	save_network(saveloc, network)
end

```

The script looks for all files in the sponges folder. Each file is provided to FlashWeave and the inferred network is then stored as a .gml file. Note that for FlashWeave performance, the sensitive and heterogeneous settings matter a lot; their importance is described in more detail in the FlashWeave publication. 

Once you have generated all the networks, feel free to load them into Cytoscape to see what they look like. 
You can also follow the R or Python tutorials to add the taxonomy to the networks. 

[Back to overview](https://ramellose.github.io/networktutorials/overview_workshop.html)

## References

[Moitinho-Silva, L., Nielsen, S., Amir, A., Gonzalez, A., Ackermann, G. L., Cerrano, C., ... & Thomas, T. (2017). The sponge microbiome project. Gigascience, 6(10), gix077.](https://academic.oup.com/gigascience/article/6/10/gix077/4082886?login=true)  

[Tackmann, J., Rodrigues, J. F. M., & von Mering, C. (2019). Rapid inference of direct interactions in large-scale ecological networks from heterogeneous microbial sequencing data. Cell systems, 9(3), 286-296.](https://www.cell.com/cell-systems/fulltext/S2405-4712(19)30271-6?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS2405471219302716%3Fshowall%3Dtrue)